---
title: "gun violence cross US"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggplot2)
library(rvest)
library(httr)
library(ggalt)
library(ggforce)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## data clean

from  1 January 2018 to 9 November 2021
```{r}
gun_violence_state_df=
read_csv("data/gun violence/mass shootings(all years).csv") %>% 
  janitor::clean_names() %>% 
  select(-incident_id,-address,-operations) %>%
  separate(incident_date,into=c("day","month","year"),sep="-") %>% 
  group_by(state,year) %>%
  mutate(number_year_killed=sum(number_killed),number_year_injured=sum(number_injured),year=as.numeric(year)) %>% 
  ungroup() %>% 
  filter(year!="17") %>% 
  mutate(year=year+2000) %>% 
  select(-city_or_county)
```

## total number of mass shooting from 2018 - 2020 in America
nationally mass shootings jumped nearly 50% during a pandemic with crippling unemployment, violent protests and idle youth.

```{r}
year_massshooting=
    gun_violence_state_df %>% 
    group_by(year) %>% 
    mutate(number_mass_shooting=n()) %>% 
    group_by(year,number_mass_shooting) %>% 
    summarise() ## result is the same as shown in the table of website of GVA(gun violence archive)

year_massshooting %>% 
  ggplot(aes(x=year,y=number_mass_shooting)) + geom_line(color="lightblue",size=1.6)+ geom_text(aes(label = number_mass_shooting)) + labs(x="Year",y="Mass Shooting",title= "Mass shooting in US")

year_massshooting %>% 
  ggplot(aes(year,number_mass_shooting)) + geom_col(color="lightblue",size=1.6,fill="lightblue",position = "stack",width=0.3,size=6) + geom_text(aes(label = number_mass_shooting),size=4,position=position_dodge(1),vjust=-0.4,angle = 0) + labs(x="Year",y="Mass Shooting",title= "Mass shooting in US",caption = "Only frist 11 months data in 2021",colour = "Cylinders",tag = "A") +geom_line(size=1,color="lightblue") + geom_point(color="blue",size=1) 

year_massshooting %>% 
  ggplot(aes(year,number_mass_shooting)) + geom_col(color="lightblue",size=1.6,fill="lightblue",position = "stack",width=0.3,size=6) + geom_text(aes(label = number_mass_shooting),size=4.5,position=position_dodge(1),vjust=-0.6,angle = 0) + labs(x="Year",y="Mass Shooting",title= "Mass shooting in US from 2018 to 2021",caption ="Source:www.gunviolencearchive.org \n (Note:only frist 11 months data in 2021) ",colour = "Cylinders",tag = "A") +geom_line(size=1,color="lightblue") + geom_point(color="blue",size=1) + 
  theme(plot.background = element_rect(colour = "lightblue", size = 3, linetype = 1, fill = "white"), plot.title = element_text(colour = "black", face = "bold", size = 20, vjust = 2.5, hjust = 0.3), plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "inches"),axis.text=element_text(colour="black")) + 
  scale_y_continuous(limits = c(0, 700))


```

## bar 



## mass shooting in US 

March 11 — WHO Declares COVID-19 a Pandemic

May 28 — US COVID-19 Deaths Pass the 100,000 Mark

June 10 — US COVID-19 Cases Reach 2 Million

January 20, 2020 CDC confirms the first U.S. laboratory-confirmed case of COVID-19 in the U.S. from samples taken on January 18 in Washington state

```{r}
year_month_state_massshooting=
  gun_violence_state_df %>%
  group_by(year,month,state) %>% 
  mutate(number_mass_shooting=n()) %>% 
  group_by(year,month,state,number_mass_shooting) %>% 
  summarise() %>% 
  mutate(year_month=str_c(year,month,sep="")) %>% 
  mutate(month=match(month,month.abb)) %>% 
  group_by(year) %>% 
  arrange(month) %>% 
  filter(year_month!="2021Nov")


year_month_state_massshooting %>% 
  mutate(year_month=fct_inorder(year_month)) %>% 
  group_by(year_month) %>% 
  count() %>% 
  rename(number_mass_shooting=n) %>% 
  ggplot(aes(year_month,number_mass_shooting)) + theme_bw(base_size=15 ) +

  geom_xspline(aes(group=1),color="lightblue",size=1.3) + geom_point(color="lightblue",size=2.5) +
  
  geom_segment(aes(x = 25, y = 10, xend = 25, yend = 32) ,size=1.6,colour = "lightblue",linetype="dashed") + 
  geom_text(aes(x=25, y= 30, label = "first case of death in US " ), colour = "red", vjust = -2, hjust = 0.5, size = 6)  + 
  

  labs(x = "Month Year ", y = "Mass Shooting", title = "National mass shooting in US ",caption="(Source:www.gunviolencearchive.org)") + theme( plot.title = element_text(colour = "black", face = "bold", 
    size = 26, vjust = 3, hjust = 0.5), plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "inches"),axis.text=element_text(colour="black",size=13),axis.title=element_text(size=20)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1))

```




## according to states,arrange total number of mass shooting from 2018 to 2021 in 48 states in an ascending way 
## bar plot
```{r}
mass_shooting_state=
  gun_violence_state_df %>% 
  group_by(state) %>%
  count() %>% 
  select(number_mass_shooting=n) %>% 
  arrange(number_mass_shooting) %>% 
  mutate(Number=number_mass_shooting)

mass_shooting_state %>% 
  ggplot(aes(x=fct_reorder(state,number_mass_shooting),y=number_mass_shooting,fill=Number)) + geom_bar(stat = "identity", colour = "black",position = "dodge") + theme_bw(base_size=15 ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1))+ labs(y= "Mass Shooting", x = "State",title="Top States with highest number of mass shooting",caption="(Source:www.gunviolencearchive.org)") +
  
  theme( plot.title = element_text(colour = "black", face = "bold", 
  size = 32, vjust = 3, hjust = 0.5), plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "inches"),axis.text=element_text(colour="black",size=6)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1,size=14),axis.text.y = element_text(size=14),axis.title=element_text(size=20)) 
```


## from 2018 to 2021, top 7 states of  mass-shooting are Illinois,California,Texas,Florida and Pennsylvania,Louisiana,New York and Ohio.

Mass shootings in California fell to 39 in 2020 from 49 the year before, other 6 states all increase in mass shooting 

```{r}




## write.csv(year_month_state_massshooting,"data/year_month_state_massshooting.csv")

year_state_massshooting=
  gun_violence_state_df %>%
    group_by(year,state) %>% 
    mutate(number_mass_shooting=n()) %>% 
    group_by(year,state,number_mass_shooting) %>% 
    summarise() 

year_state_massshooting %>% 
  filter(state%in%c("Illinois","California","Texas","Florida","Pennsylvania","Louisiana","New York","Ohio")) %>% 
  ggplot(aes(x=year,y=number_mass_shooting,color=state)) +  +  facet_grid(~fct_reorder(state,number_mass_shooting))+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1))

year_state_massshooting %>% 
  filter(state%in%c("Illinois","California","Texas","Florida","Pennsylvania","Louisiana","New York","Ohio")) %>% 
  ggplot(aes(x=year,y=number_mass_shooting,fill=state)) + geom_col(size=1.3) +  facet_grid(~fct_reorder(state,number_mass_shooting))+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1)) + labs(title="Mass shooting in top 8 states from 2018 to 2021",caption="Source:www.gunviolencearchive.org") + theme(plot.background = element_rect(colour = "lightblue", size = 3, linetype = 1, 
  fill = "white"), plot.title = element_text(colour = "black", face = "bold", 
  size = 14, vjust = 3, hjust = 0.5), plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "inches"),axis.text=element_text(colour="black",size=10))

```

```{r}
year_state_injured_killed=
  gun_violence_state_df %>%
  group_by(state,year) %>% 
  mutate(Victims=number_year_killed + number_year_injured,Year=year) %>% 
  select(Year,state,number_year_killed,number_year_injured,Victims) %>% 
  distinct()

  
year_state_injured_killed %>% 
  filter(state%in%c("Illinois","California","Texas","Florida","Pennsylvania","Louisiana","New York","Ohio")) %>% 
  ggplot(aes(Year,Victims,fill=state)) + geom_col(width=0.4) + labs(title="Victims in top 8 states of US",caption ="Source:www.gunviolencearchive.org") +
  theme(plot.background = element_rect(colour = "lightblue", size = 3, linetype = 1, fill = "white"), plot.title = element_text(colour = "black", face = "bold", size = 20, vjust = 2.5, hjust = 0.3), plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "inches"),axis.text=element_text(colour="black")) 
  
  
  
  
```





## number of people killed  and injured in mass shooting from  Juanary 2018 to 2021 (since 2021 has not ended,so the latest date is 2021.11.9)
```{r}
gun_violence_state_df %>% 
  group_by(year) %>% 
  mutate(year_nation_killed=sum(number_year_killed),year_nation_injured=sum(number_year_injured)) %>% 
  group_by(year,year_nation_killed,year_nation_injured) %>% 
  summarise()
```



```{r}
plot_usmap(include = c("NY")) +
  labs(title = "Western US States",
       subtitle = "These are the states in the Pacific Timezone.")
```

