---
title: "heatmap"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(gridExtra)
library(leaflet)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE}
historic_NYPD_df = read_csv("./data/NYPD_Shooting_Incident_Data__Historic_.csv") %>%
  separate('OCCUR_DATE', into=c("month","day","year"),sep = "/") %>%
  filter(year>=2018)
year_to_date_df = read_csv("./data/NYPD_Shooting_Incident_Data__Year_To_Date_.csv")%>%
  separate('OCCUR_DATE', into=c("month","day","year"),sep = "/")
df = bind_rows(historic_NYPD_df,year_to_date_df)
```


### The change of crime frequency  
It’s possible that the crime drequency will have changed during the pandemic. We therefore explored whether the crime by time of day changed pre-pandemic versus during the pandemic.\
To accomplish this, we created a heat map of gun crime frequency by time of day pre (from 2018 to 2019) and during COVID (from 2020 to 2021).
```{r, warning=FALSE}

df_period_before = df %>%
  arrange(year,month,day)%>%
  filter(year <= 2019) %>% 
  separate(OCCUR_TIME,into = c("hour","minute","second"),sep = ":")%>%
  mutate(hour=as.numeric(hour))%>%
  group_by(hour) %>%
  summarise(count=n(),
            time="2018-2019")

df_period_after = df %>%
  arrange(year,month,day)%>%
  filter(year>= 2020) %>% 
  separate(OCCUR_TIME,into = c("hour","minute","second"),sep = ":")%>%
  mutate(hour=as.numeric(hour)) %>% 
  group_by(hour) %>%
  summarise(count=n(),
            time="2020-2021")

df_period =
  bind_rows(df_period_before,df_period_after) %>% 
  plot_ly(
    x = ~hour, y = ~time, z = ~ count, type = "heatmap", colors = "YlGn"
  ) %>%
  colorbar(title = "Crimes Number", x = 1.1, y = 0.8)

layout(df_period, title = "Crime frequency during a day", 
       xaxis = list(title = "Hour", tick0=0, dtick=6), 
       yaxis = list(title = ' '), 
       width = 850, height = 400,
       annotations = 
 list(x = 1.15, y = -0.15, text = "Source: https://opendata.cityofnewyork.us", 
      showarrow = F, xref='paper', yref='paper', 
      xanchor='right', yanchor='auto', xshift=0, yshift=0,
      font=list(colour = "Cylinders",tag = "A")))

ggsave("heatmap of hours.png")
```
*Comments:* \
Based on the heat map,  it is easier to obtain that gun crime events most happened in the afternoon and evening. Since people are more convenient to commit a crime after dark.\
We can also that gun crime events most happened in the afternoon and evening. Since people are more convenient to commit a crime after dark. As seen in this plot, there begin  to commit a crime in the afternoon rush hour times of 3-6 PM during COVID as compared to pre-COVID. In addition, there was more gun crimes during COVID.



### Location of gun crime 
This map displays the geolocation of gun crime in New York City from 2018 to 2019. The data are interactive and organized by borough.
```{r, warning=FALSE}
data_pre = 
  df %>% 
  janitor::clean_names() %>% 
  arrange(year,month,day)%>%
  filter(year <= 2019) %>%
  rename(long = longitude, lat = latitude) %>% 
  group_by(boro) %>% 
  mutate(crime_number = n(),
         crime_number = as.numeric(crime_number))

pal = colorNumeric(
  palette = "Reds",
  domain =  c(0:1300),
  na.color = "white")

data_pre %>% 
  mutate(
    label = str_c("<b>crime number: ", crime_number, "</b><br>location: ", lon_lat , sep = "") ) %>%
  leaflet() %>% 
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircles(lng = ~long, lat = ~lat, weight = 5, stroke = FALSE, radius = ~sqrt(crime_number)*6, popup = ~ label, color = ~pal(crime_number))%>% 
  addMarkers(clusterOptions = markerClusterOptions(), popup = ~ label) %>% 
  addLegend("topright", pal = pal, values = ~crime_number,
    title = "2019-2020 Total Gun Crime",
    opacity = 1
  ) %>% 
   setView(-73.8399986, 40.746739, zoom = 10.5)





```

This map displays the geolocation of gun crime in New York City from 2020 to 2021. The data are interactive and organized by borough.
```{r, warning=FALSE}
data_during = 
  df %>% 
  janitor::clean_names() %>% 
  arrange(year,month,day)%>%
  filter(year >= 2020) %>%
  rename(long = longitude, lat = latitude) %>% 
  group_by(boro) %>% 
  mutate(crime_number = n(),
         crime_number = as.numeric(crime_number))

pal = colorNumeric(
  palette = "Reds",
  domain =  c(0:1300),
  na.color = "white")

data_during %>% 
  mutate(
    label = str_c("<b>crime number: ", crime_number, "</b><br>location: ", lon_lat , sep = "") ) %>%
  leaflet() %>% 
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircles(lng = ~long, lat = ~lat, weight = 5, stroke = FALSE, radius = ~sqrt(crime_number)*6, popup = ~ label, color = ~pal(crime_number))%>% 
  addMarkers(clusterOptions = markerClusterOptions(), popup = ~ label) %>% 
  addLegend("topright", pal = pal, values = ~crime_number,
    title = "2020-2021 Total Gun Crime",
    opacity = 1
  ) %>% 
 setView(-73.8399986, 40.746739, zoom = 10.5)
```

#### Year to Year Comparison

There appear to some incremental changes in the geolocation of gun crime from 2018 to 2021, including an increase in clustering in the all the region during covid period(2020-2021), relative to pre covid(2018-2019). Further, more gun crime happened during the covid in all neighborhoods of newyork city.\

#### Neighborhood Clustering

There is a rapid increase of gun crime in Bronx and Queens during covid. 
