---
title: "A logistic regression approach on the malicity of shooting incident: a murder or a homicide?"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{css, echo=FALSE}
.scroll-200 {
  max-height: 200px;
  overflow-y: auto;
  background-color: inherit;
}
```

```{r setup, include=FALSE}
library(tidyverse)
library(glmnet)
library(pROC)
library(RTextTools)
```
In the NYPD shooting incident database, a variable is called STATISTICAL_MURDER_FLAG which indicates if this shooting incident is classified as a murder. It has a binary outcome "TRUE" or "FALSE". We will use logistic regression model to identify if a shooting can be classified as a murder.

## Data Import
At first, import data from the NYPD shooting incident database
```{r echo=TRUE, message=FALSE, warning=FALSE}
historic_NYPD_df = read_csv("./data/NYPD_Shooting_Incident_Data__Historic_.csv") 
year_to_date_df = read_csv("./data/NYPD_Shooting_Incident_Data__Year_To_Date_.csv")
df = bind_rows(historic_NYPD_df,year_to_date_df) 
 df = df %>%
  separate(OCCUR_TIME,into = c("hour","minute","second"),sep = ":")%>%
  separate(OCCUR_DATE,into = c("month","date","year"),sep = "/")%>%
  mutate(hour=as.numeric(hour),
    time_period = case_when(
    hour>=0 & hour<6 ~ "MIDNIGHT",
           hour>=6 & hour<12 ~ "MORNING",
           hour>=12 & hour<18 ~ "AFTERNOON",
           hour>=18 & hour<=23 ~ "EVENING"),
         month = as.factor(month))%>%
  select(BORO,STATISTICAL_MURDER_FLAG,PERP_AGE_GROUP,PERP_SEX,PERP_RACE,VIC_AGE_GROUP,VIC_SEX,VIC_RACE,time_period,month)%>%
  mutate(STATISTICAL_MURDER_FLAG = ifelse(STATISTICAL_MURDER_FLAG=="TRUE", 1, 0),
         BORO = as.factor(BORO),
         PERP_AGE_GROUP = as.factor(PERP_AGE_GROUP),
         PERP_SEX = as.factor(PERP_SEX),
         PERP_RACE = as.factor(PERP_RACE),
         VIC_AGE_GROUP = as.factor(VIC_AGE_GROUP),
         VIC_SEX = as.factor(VIC_SEX),
         VIC_RACE = as.factor(VIC_RACE))%>%
  na.omit()
```
there are some irregular value and we need to deal with them.
```{r}
df=df[df$PERP_AGE_GROUP != "UNKNOWN", ]
df=df[df$PERP_AGE_GROUP != "1020", ]
df=df[df$PERP_AGE_GROUP != "940", ]
df=df[df$PERP_AGE_GROUP != "224", ]
df=df[df$PERP_SEX != "U", ]
df=df[df$PERP_RACE != "UNKNOWN", ]
df=df[df$VIC_SEX != "U", ]
df=df[df$VIC_AGE_GROUP != "UNKNOWN", ]
```


## Model Selection
First, build the initial logistic regression model with all variables

```{r, class.output="scroll-200"}

fit = glm(STATISTICAL_MURDER_FLAG ~ BORO+PERP_AGE_GROUP+PERP_SEX+PERP_RACE+VIC_AGE_GROUP+VIC_SEX+VIC_RACE+time_period+month,data=df,family = binomial())
summary(fit)
```
we can see from the above result that some variables do not pass the significance test. Now, we can use the stepwise method to choose variables.



```{r, class.output="scroll-200"}

step(fit)
```

Now, we test the reduced model from the above result


```{r,class.output="scroll-200"}

fit_2 = glm(formula = STATISTICAL_MURDER_FLAG ~ BORO + PERP_AGE_GROUP + 
    PERP_RACE + VIC_AGE_GROUP + VIC_RACE + time_period, family = binomial(), 
    data = df)
summary(fit_2)
```


We can see that PERP_RACE and VIC_RACE still don't pass the significance test, now we drop these two and run again
```{r,,class.output="scroll-200"}

fit_3 = glm(formula = STATISTICAL_MURDER_FLAG ~ BORO + PERP_AGE_GROUP + 
              VIC_AGE_GROUP +  time_period, family = binomial(), 
    data = df)
summary(fit_3)
```

From the above result, we use dummy variables for BOROQUEENS,BOROSTATEN ISLAND,PERP_AGE_GROUP18-24
```{r}
df$QUEENS = ifelse(df$BORO=="QUEENS",1,0)
df$STATEN = ifelse(df$BORO=="STATEN ISLAND",1,0)
df$BROOKLYN = ifelse(df$BORO=="BROOKLYN",1,0)
df$MANHATTAN = ifelse(df$BORO=="MANHATTAN",1,0)
df$PERP_18 =ifelse(df$PERP_AGE_GROUP=="18-24",1,0)
df$PERP_25 =ifelse(df$PERP_AGE_GROUP=="25-44",1,0)
df$PERP_45 =ifelse(df$PERP_AGE_GROUP=="45-64",1,0)
df$PERP_65 =ifelse(df$PERP_AGE_GROUP=="65+",1,0)
```

Run logistic regression with these dummy variables again and use stepwise method to filter.


```{r,class.output="scroll-200"}

fit_4 = glm(formula = STATISTICAL_MURDER_FLAG ~ QUEENS + STATEN + BROOKLYN + MANHATTAN + PERP_18 + PERP_25 + PERP_45 + PERP_65 + 
              VIC_AGE_GROUP +  time_period, family = binomial(), 
    data = df)
summary(fit_4)
step(fit_4)
```

now,we can run the model for the last time


```{r,class.output="scroll-200"}

fit_5 = glm(formula = STATISTICAL_MURDER_FLAG ~ BROOKLYN + MANHATTAN + 
    PERP_25 + PERP_45 + PERP_65 + VIC_AGE_GROUP + time_period, 
    family = binomial(), data = df)
summary(fit_5)
```

This model seems good with all variables having significance. Now we test for the accuracy 
```{r}
ifMudrer = function(glm.model, threshold = 0.5) {
  pre.pro = predict(glm.model, type = "response")
  pre = if_else(pre.pro >= threshold, 1, 0)
}
recall_accuracy(df$STATISTICAL_MURDER_FLAG,ifMudrer(fit_5))
```


## ROC plot
A receiver operating characteristic curve, or ROC curve, is a graphical plot that illustrates the diagnostic ability of a binary classifier system as its discrimination threshold is varied. The ROC curve is created by plotting the true positive rate (TPR) against the false positive rate (FPR) at various threshold settings.
```{r message=FALSE}
roc(df$STATISTICAL_MURDER_FLAG, predict(fit_5, type = "response"), algorithm = 3) %>% 
  plot(auc.polygon = T,  print.thres = T, print.auc = T)
```

The area under the curve can be used to quantitatively measure the effect of the model. The area under the ROC curve is denoted as AUC, so the larger AUC is, the better the model is. AUC for our model is 0.575 which is acceptable.


## Conclusion
At first, we would like to see the significance for each variable
```{r}
exp(fit_5$coefficients)
```
With the above results, we can have the following conclusion:  
-Shooting incidents happen in BROOKLYN and MANHATTAN are less likely to be a murder.  
-When perpetrators' age increases, the probability that this shooting incident to become a murder also increases. This conclusion holds for victims.  
-Shooting incidents happen in the morning are more like to be a murder.  

## Prediction
For example, we have two shooting incidents now. We will use this model to predict which one has a higher probability to be a murder.  
--A 36-year-old white female was killed by a 20-year-old black male at 10pm in Brooklyn.  
--A 48-year-old Asian male was killed by a 54-year-old white male at 10am in Queens.  
```{r warning=FALSE}
test_data = data_frame(BROOKLYN = c(1,0) , MANHATTAN = c(0,0) ,PERP_25 = c(1,0),  PERP_45 =  c(0,1), PERP_65 = c(0,0) , VIC_AGE_GROUP = c("25-44","45-64"),VIC_RACE = c("WHITE","ASIAN"),VIC_SEX = c("F","M"),time_period=c("EVENING","MORNING"))
test_data
test_data$prob = predict(fit_5, newdata=test_data, type="response")
test_data$prob

```
From the prediction, we can see that the second one has a higher possibility to be a murder.
